Learning Actionable Analytics (with applications for reducing defects and reducing runtimes)

### 2015-05-02 manuscript.tex
Updates from ShareLaTeX

\begin{document}
  \title{Learning Actionable Analytics 
 (with Applications for Reducing Defects and Reducing Runtimes)}

@@

 This paper uses  HOW to learn methods
 to reduce defects and decrease program runtime.
 For the test data used in this paper,
 we have observed that the improvements generated by HERE can reduce
 defect prediction and runtimes by up to
 50\% of their original values.
  \end{abstract}
  \begin{IEEEkeywords}
Instance-based reasoning, defect prediction, random forest, optimization.
  \end{IEEEkeywords}

\section{Introduction}
In the $21^{st}$ century, it is now impossible to manually browse all
the available software defect data.  For example, as of April 2015,
Mozilla Firefox has over a million bug reports on software projects.  While it is
now possible to automatically analyze such data with data miners, at
some stage a group of business users will have to convene to {\em
  interpret the results} (e.g., to decide if it is wise to deploy the
results as a defect reduction method within an organization).

These business  users are now demanding that data mining tools
be augmented with tools to support  business-level
interpretation of that data. For example,
at a  panel on software analytics at ICSE'12,
industrial practitioners lamented the state of the art in data mining
and software engineering~\cite{menzies12a}. Panelists commented that
``prediction is all well and good, but what about decision
making?''. That is, these panelists are more interested in the interpretations
that follow the mining, rather than just  the mining. For example:
\bi
\item Instead of just accepting  {\em predictions} on how many 
 software defects
to expect,  business users now demand {\em policies} that
reduce the likelihood of those defects;
\item Instead of just accepting {\em predictions} on the runtime
time of their software, business users might now demand
{\em policies} that reduce that runtime.
\ei
In order to satisfy the demands of this new kind of business user, we  explore the generation of {\em policies} in the context
of software analytics. Such {\em policies} are recommendations of what
to change in order to better achieve some goal. 
Note that such {\em policy generation} is very different to the standard
data mining tasks of prediction some discrete or numeric class.
Predictors
and classifiers only tell us what {\em is} while policy 
generators tell us {\em what to change}. For the kind of business user
discussed above, policy generators are the prefeered tool since,
as one gruff user once shouted at us ``don't tell me what \underline{is};
tell me what to \underline{change}''.

This paper presents the following general framework for extending  predictor systems
such that they become 
policy generators:
\begin{small}
\[
project\;data \rightarrow \left\{\begin{array}{l}\mathit{train}
\rightarrow \left\{\begin{array}{l}\mathit{predictor\;(e.g.\;CART\;or\; RF)}\\
\mathit{policy\; generator\; (e.g.\; HOW)}\end{array}\right.
\\
~\\
\mathit{test}  \rightarrow  \left\{\begin{array}{r@{~}l}
\mathit{before}& =\mathit{predictor}(\mathit{test})\\
\mathit{after}& =\mathit{predictor}(\; \mathit{policy}(\mathit{test})\;)\\
\Delta& =\mathit{after} / \mathit{before}
\end{array}\right.
\end{array} \right.
\]
\end{small}
In the above:
\bi
\item
Project data is divided into a train and test set.
\item
From the train set, we learn {\em both} a policy generator as well
as a predictor using (e.g.)
CART (for making numeric predictions~\cite{Breiman1984}) or Random Forests (for making predictions
about symbolic classes~\cite{Breiman2001}). 
\item
The test data is then studied to generate two predictions:
a {\em before} prediction (on the raw test data); then an {\em after}
prediction (after applying the policies learned from the training data).
\item
The improvement delta $\Delta$ is then computed by dividing {\em after} and {\em before}.
\ei
This paper presents a novel policy generator called ``HOW''.
HOW is a near-linear time  instance-based
local search algorithm that looks for improvement trends in the data,
then adjusting examples to nudge them along that trend.
The experiments of this paper apply HOW to learning policies for 
software defect reduction and for compiling software that runs faster.
At least for the data explored here, HOW's reductions in defects and runtimes
are very large-- up to 50\% fewer defects and up to 50\% faster runtimes.

>>>>

\begin{document}
  \title{Learning Actionable Analytics 
 (with applications for reducing defects and reducing runtimes)}

@@

 This paper uses  HOW to learn methods
 to reduce defects and decrease program runtime.
 For the test data of this paper, the improvements generated by HERE can reduce
 defect counts as well as runtimes by up to
 50\% of their original values.
  \end{abstract}
  \begin{IEEEkeywords}
Instance-based reasoning, defect prediction, random forest, optimization.
  \end{IEEEkeywords}

\section{Introduction}
Business  users are now demanding that data mining tools
be augmented with tools to support  business-level
interpretation of that data. For example,
at a  panel on software analytics at ICSE'12,
industrial practitioners lamented the state of the art in data mining
and software engineering~\cite{menzies12a}. Panelists commented that
``prediction is all well and good, but what about decision
making?''. That is, these panelists are more interested in the interpretations
that follow the mining, rather than just  the mining. For example:
\bi
\item Instead of just accepting  {\em predictions} on how many 
 software defects
to expect,  business users might now demand a {\em plan} to
reduce the likelihood of those defects;
\item Instead of just accepting {\em predictions} on the runtime
time of their software, business users might now demand
a {\em plan} to reduce that runtime.
\ei
In these two points, we say that a ``plan'' is a  set of proposed changes in order to better achieve some goal. 
Generating {\em plans} is a  very different task the standard
data mining tasks of {\em predicting} some discrete or numeric class.
Predictors
and classifiers only tell us what {\em is} while planners 
  tell us {\em what to change}. For the kind of business user
discussed above, planners are the preferred to mere prediction systems
since, as one gruff user once shouted at us:
\begin{quote}
{\em ``Don't tell me what \underline{is};
tell me what to \underline{change}.''}
\end{quote}
In order to satisfy the demands of this new kind of business user, we  study the generation of {\em plans} in the context
of software analytics.  
As reported in this paper, that study makes five contributions:
\be
\item A general framework for augmenting a prediction system such that
it becomes a planning system;
\item A  test that recognizes  when we should {\em not}  plan: in summary, our kind of planning is
not recommended in domains that  generate poorly performing predictors;
\item A general verification rig for assessing the value of plans generated within this rig.
\item A novel near-linear-time instance-based planner called ``HOW''. 
\item An experimental validation that the HOW planner is useful for reducing the
expected values of 
\bi
\item Software defects in certain object-oriented JAVA systems listed in \tion{oo}:  Ant Camel, Ivy, Jedit, Log4, Lucene, Poi, Synapse, Velocity, Xalan;
\item The runtime of the software systems discussed in \tion{cpu}: Berkeley DB, Apache, SQLite, LLVM, 
  x264.
\ei
\ei
HOW is an {\em instance-based planner} in that it make its conclusions by
studying the space of examples. This is a very different approach to
{\em model-based planners} that use the model
as sub-routine within a Monte Carlo~\cite{me07f} or evolutionary 
programming rig~\cite{krall14,harman12dec}. 
We prefer HOW's instance-based approach since
it is not dependent on any underlying model.
This is important since:
\bi
\item
When some models are hard to access or slow to run (see examples, below),
then it useful to reason from historical logs rather than from hard-to-collect
model output.
\item
For reasons of confidentionality or priorpiarty restrictions, 
It can be harder to access 
\item
Building and certifying a model   can take years. For example, in previous work~\cite{me07f} we used models
developed by Boehm's group at the University of Southern California.
Some of those models took decades to develop and mature (from 1981~\cite{boehm81} to 2000~\cite{boehm00b}). 
\ei
The rest of this paepr is structued as follows.

\section{From ``Prediction'' to ``Planning''}

approach if the underlying model

This paper presents the following general framework for extending  predictor systems
such that they become 
policy generators:

\begin{small}
\[
\begin{array}{r}
project\\
data
\end{array}\rightarrow \left\{\begin{array}{l}\mathit{train}
\rightarrow \left\{\begin{array}{l}\mathit{build\;predictor\;(e.g.\;CART\;or\; RF)}\\
\mathit{build\;planner\; (e.g.\; HOW)}\end{array}\right.
\\
~\\
\mathit{test}  \rightarrow  \left\{\begin{array}{l@{~}l}
\mathit{before}& =\mathit{predictor}(\mathit{test})\\
\mathit{if\; before} & \ge  \mathif{satisfactory\; then}\\
 &
\begin{array}{l@{~}l}

\mathit{test'}& = \mathit{planner}(\mathit{test})\\
\mathit{after}& =\mathit{predictor}(\mathit{test'})\\ 
\Delta& =\mathit{after} / \mathit{before}
\end{array}
\end{array}\right.
\end{array} \right.
\]
\end{small}
In the above:
\bi
\item
Project data is divided into a train and test set.
\item
From the train set, we learn {\em both} a policy generator as well
as a predictor using (e.g.)
CART (for making numeric predictions~\cite{Breiman1984}) or Random Forests (for making predictions
about symbolic classes~\cite{Breiman2001}). 
\item
The test data is then studied to generate two predictions:
a {\em before} prediction (on the raw test data); then an {\em after}
prediction (after applying the policies learned from the training data).
\item
The improvement delta $\Delta$ is then computed by dividing {\em after} and {\em before}.
\ei
This paper presents a novel policy generator called ``HOW''.
HOW is a near-linear time  instance-based
local search algorithm that looks for improvement trends in the data,
then adjusting examples to nudge them along that trend.
The experiments of this paper apply HOW to learning policies for 
software defect reduction and for compiling software that runs faster.
At least for the data explored here, HOW's reductions in defects and runtimes
are very large-- up to 50\% fewer defects and up to 50\% faster runtimes.
